One-arm spirals in locally isothermal, radially structured self-gravitating discs

### 2014-11-06 paper/paper.tex
minor edits

\maketitle
\begin{abstract} 
  We present numerical simulations of astrophysical discs with an 
  imposed temperature profile and in which a sub-portion of it is
  self-gravitating. This might represent, for example, mass  
  built-up in a circumstellar or protoplanetary
  disc at the outer edge of a `dead zone', in which there
  is inefficient magneto-hydrodynamic turbulence to enable accretion. 
  We conduct numerical gravito-hydrodynamic experiments in 2D using
  the FARGO code and in 3D with the ZEUS-MP and PLUTO codes. In such
  discs we observe the development of a 
  spiral mode with azimuthal wavenumber $m=1$ with co-rotation radius 
  outside the bulk of the spiral arm, resulting in a nearly stationary 
  one-arm spiral pattern in the self-gravitating region of the
  disc. The development of this one-arm 
  spiral does not require the movement of  
  the central star, as found in previous studies. Instead, we find it 
  to be associated with the imposed 
  temperature gradient and the radial disc structure. 
  We apply results from existing linear theory and attribute the
  growth of this spiral to the non-conservation of angular momentum
  for non-axisymmetric linear perturbations in discs with a forced temperature
  gradient. Such interaction may lead to an unstable exchange of angular momentum between the spiral 
  and the background disc. However, because the exchange mechanism does
  not explicitly depend on self-gravity, we speculate it may also
  enhance low-frequency one-arm spirals in non-self-gravitating
  discs.

>>>>

\maketitle
\begin{abstract} 
  We present numerical simulations of astrophysical discs with an 
  imposed temperature profile and in which a sub-portion of it is
  self-gravitating. This might represent, for example, mass  
  built-up in a circumstellar or protoplanetary
  disc at the outer edge of a `dead zone', in which there
  is inefficient magneto-hydrodynamic turbulence to enable accretion. 
  We conduct numerical gravito-hydrodynamic experiments in 2D using
  the FARGO code and in 3D with the ZEUS-MP and PLUTO codes. In such
  discs we observe the development of a 
  spiral mode with azimuthal wavenumber $m=1$ with co-rotation radius 
  outside the bulk of the spiral arm, resulting in a nearly stationary 
  one-arm spiral pattern in the self-gravitating region of the
  disc. The development of this one-arm 
  spiral does not require the movement of  
  the central star, as found in previous studies. Instead, we find it 
  to be associated with the imposed 
  temperature gradient and the radial disc structure. 
  We apply results from existing linear theory and attribute the
  growth of this spiral to the non-conservation of angular momentum
  for non-axisymmetric linear perturbations in discs with a forced temperature
  gradient. This may lead to an unstable exchange of angular momentum between the spiral 
  and the background disc. However, because the interaction does
  not explicitly depend on self-gravity, we speculate it may also
  enhance low-frequency one-arm spirals in non-self-gravitating
  discs.
